NGO analysis

### 2022-08-01 Dashboard/report.Rmd
Tidying. Changing to is.ggplot() test.

## Function to read html file --------------------------------------------------
read_proff <- function(file, start, end) {
  # Specifying encoding to deal with occasional errors
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    # creating a currency code column and turning current currency codes into
    # NAs
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    # filling the currency column
    fill(
      valutakode,
      .direction = "updown"
    ) |>
    # removing dates and redundant currency
    filter(
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate(
      # removing white space in numbers
      value = str_replace_all(value, "\\s", ""),
      # turning numbers into true numbers
      value = as.numeric(value),
      # removing values that are years
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    # removing any remaining duplicate rows
    distinct() |>
    # pivoting to create 'tidy' variables
    pivot_wider(
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    # tidying names
    clean_names() |>
    # removing headings from the table
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |>
    # making it clear what lonn refers to
    rename(
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # arranging by year for practical display
    arrange(desc(year))
}

@@

## Functions to read Excel file ------------------------------------------------
### First Excel transformation -------------------------------------------------
# creating df used by NGO title function and maximum scale function
widen_excel_test <- function(file) {
  # Reading Excel file
  read_excel(file) |>
    # cleaning names to enable later operations
    clean_names()
}

@@

# Creating general warning
general_warning <- if (datatest == TRUE ||
  sum(!is.na(htmldata$arsresultat)) == 0 ||
  sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0 ||
  (sum(!is.na(htmldata$lederlonn)) == 0 &&
    sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0)) {
  "## Note:"
} else {
  ""
}
```

@@

```{r balance_plot, fig.cap="Based on the entry \`Sum egenkapital og gjeld' in the Proff.no data."}
## Balance--------------------------------------------------------------------
balance_plot <-
  # testing for the presence of data
  if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0) {
    # Printing 'no information' plot if no data
    NA
  } else {
    htmldata |>
      # rename for plotly hover
      rename(
        Year = year,
        `Total equity/liabilities` = sum_egenkapital_og_gjeld,
        `Positive equity` = balance_positive
      ) |>
      ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale +
      # using a function to ensure uniform colours irrespective of data
      scale_fill_green(htmldata$balance_positive) +
      labs(
        x = "", y = "",
        caption = currency_caption, title = "Total equity/liabilities (in thousands)"
      ) +
      theme_light() +
      theme(legend.position = "none")
  }

# Printing the plot if the data are available
if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) != 0) {
  balance_plot
}
```

```{r revenue_expenses_plot, fig.asp=1, out.width="50%", fig.cap="Revenue and Expenses", fig.subcap=c("Based on the category \`Revenue' in the Excel sheet.", "Based on the category \`Expenses' in the Excel sheet.")}
## Revenue--------------------------------------------------------------------
revenue_plot <-
  # testing for data quality
  if (datatest == TRUE) {
    # printing 'bad data' plot if data are bad
    NA
  } else {
    exceldata |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        `Private contributions` = revenue_private_contributions,
        `Membership fees` = revenue_membership_fees,
        `Public grants` = revenue_public_grants,
        Investments = revenue_investments,
        `Operational activities` = revenue_operational_activities,
        `Commercial activities` = revenue_commercial_activities,
        Other = revenue_other
      ) |>
      # Pivoting to use fill
      pivot_longer(
        c(`Private contributions`:Other),
        names_to = "Revenue source"
      ) |>
      # Ordering for the plot
      mutate(
        `Revenue source` =
          ordered(`Revenue source`,
            levels =
              c(
                "Other", "Commercial activities",
                "Operational activities", "Investments",
                "Membership fees", "Private contributions",
                "Public grants"
              )
          )
      ) |>
      ggplot(aes(Year, value, fill = `Revenue source`)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum)
      ) +
      # Specifying the number of values, as the base palette only includes 6
      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
      labs(x = "", y = "", title = "Revenue (in thousands)") +
      theme_light() +
      theme(title = element_text(size = 14))
  }

# Printing the plot if the data are OK
if (datatest == FALSE && excel_zero_check("Revenue") != 0) {
  revenue_plot
}

## Expenses -------------------------------------------------------------------
expenses_plot <-
  # testing for data quality
  if (datatest == TRUE) {
    NA
  } else {
    exceldata |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        Other = expenses_other,
        Administrative = expenses_administrative,
        Fundraising = expenses_fundraising,
        `Program services` = expenses_program_services
      ) |>
      # Pivoting to use fill
      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
      # Ordering for the plot
      mutate(
        Expenses = ordered(Expenses,
          levels =
            c("Program services", "Other", "Fundraising", "Administrative")
        )
      ) |>
      ggplot(aes(Year, value, fill = Expenses)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum)
      ) +
      scale_fill_met_d("Hokusai2") +
      labs(x = "", y = "", title = "Expenses (in thousands)") +
      theme_light() +
      theme(title = element_text(size = 14))
  }

# Printing the plot if the data are OK
if (datatest == FALSE && excel_zero_check("Expenses") != 0) {
  expenses_plot
}
```

@@

# Printing the plot if the data are available
if (sum(!is.na(htmldata$lederlonn)) != 0 ||
  sum(!is.na(htmldata$leder_annen_godtgjorelse)) != 0) {
  CEO_plot
}
```

>>>>

## Function to read html file --------------------------------------------------
read_proff <- function(file, start, end) {
  # Specifying encoding to deal with occasional errors
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    # filling the currency column
    fill(
      valutakode,
      .direction = "updown"
    ) |>
    # removing dates and redundant currency
    filter(
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate(
      # removing white space in numbers
      value = str_replace_all(value, "\\s", ""),
      # turning numbers into true numbers
      value = as.numeric(value),
      # removing values that are years
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    # removing any remaining duplicate rows
    distinct() |>
    # pivoting to create 'tidy' variables
    pivot_wider(
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    # tidying names
    clean_names() |>
    # removing headings from the table
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |>
    # making it clear what lonn refers to
    rename(
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # arranging by year for practical display
    arrange(desc(year))
}

@@

## Functions to read Excel file ------------------------------------------------
### First Excel transformation -------------------------------------------------
# creating df used by NGO title function, maximum scale function, and excel zero
# check
widen_excel_test <- function(file) {
  # Reading Excel file
  read_excel(file) |>
    # cleaning names to enable later operations
    clean_names()
}

@@

# Creating general warning
general_warning <- if (
  datatest == TRUE ||
    sum(!is.na(htmldata$arsresultat)) == 0 ||
    sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0 ||
    (sum(!is.na(htmldata$lederlonn)) == 0 &&
      sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) ||
    excel_zero_check("Revenue") == 0 ||
    excel_zero_check("Expenses") == 0
) {
  "## Note:"
} else {
  ""
}
```

@@

```{r balance_plot, fig.cap="Based on the entry \`Sum egenkapital og gjeld' in the Proff.no data."}
## Balance--------------------------------------------------------------------
balance_plot <-
  # testing for the presence of data
  if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0) {
    NA
  } else {
    htmldata |>
      # rename for plotly hover
      rename(
        Year = year,
        `Total equity/liabilities` = sum_egenkapital_og_gjeld,
        `Positive equity` = balance_positive
      ) |>
      ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale +
      # using a function to ensure uniform colours irrespective of data
      scale_fill_green(htmldata$balance_positive) +
      labs(
        x = "", y = "",
        caption = currency_caption, title = "Total equity/liabilities (in thousands)"
      ) +
      theme_light() +
      theme(legend.position = "none")
  }

# Printing the plot if the data are available
if (is.ggplot(balance_plot)) {
  balance_plot
}
```

```{r revenue_expenses_plot, fig.asp=1, out.width="50%", fig.cap="Revenue and Expenses", fig.subcap=c("Based on the category \`Revenue' in the Excel sheet.", "Based on the category \`Expenses' in the Excel sheet.")}
## Revenue--------------------------------------------------------------------
revenue_plot <-
  # testing for data quality
  if (datatest == TRUE || excel_zero_check("Revenue") == 0) {
    NA
  } else {
    exceldata |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        `Private contributions` = revenue_private_contributions,
        `Membership fees` = revenue_membership_fees,
        `Public grants` = revenue_public_grants,
        Investments = revenue_investments,
        `Operational activities` = revenue_operational_activities,
        `Commercial activities` = revenue_commercial_activities,
        Other = revenue_other
      ) |>
      # Pivoting to use fill
      pivot_longer(
        c(`Private contributions`:Other),
        names_to = "Revenue source"
      ) |>
      # Ordering for the plot
      mutate(
        `Revenue source` =
          ordered(`Revenue source`,
            levels =
              c(
                "Other", "Commercial activities",
                "Operational activities", "Investments",
                "Membership fees", "Private contributions",
                "Public grants"
              )
          )
      ) |>
      ggplot(aes(Year, value, fill = `Revenue source`)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum)
      ) +
      # Specifying the number of values, as the base palette only includes 6
      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
      labs(x = "", y = "", title = "Revenue (in thousands)") +
      theme_light() +
      theme(title = element_text(size = 14))
  }

# Printing the plot if the data are OK
if (is.ggplot(revenue_plot)) {
  revenue_plot
}

## Expenses -------------------------------------------------------------------
expenses_plot <-
  # testing for data quality
  if (datatest == TRUE || excel_zero_check("Expenses") == 0) {
    NA
  } else {
    exceldata |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        Other = expenses_other,
        Administrative = expenses_administrative,
        Fundraising = expenses_fundraising,
        `Program services` = expenses_program_services
      ) |>
      # Pivoting to use fill
      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
      # Ordering for the plot
      mutate(
        Expenses = ordered(Expenses,
          levels =
            c("Program services", "Other", "Fundraising", "Administrative")
        )
      ) |>
      ggplot(aes(Year, value, fill = Expenses)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum)
      ) +
      scale_fill_met_d("Hokusai2") +
      labs(x = "", y = "", title = "Expenses (in thousands)") +
      theme_light() +
      theme(title = element_text(size = 14))
  }

# Printing the plot if the data are OK
if (is.ggplot(expenses_plot)) {
  expenses_plot
}
```

@@

# Printing the plot if the data are available
if (is.ggplot(CEO_plot)) {
  CEO_plot
}
```
